---
title: "Abundance Survey Summary Statistics, Revisions -- Temperature?"
output: html_notebook
---
# Intro
10/30/2017
This is an addition to Summary statistics 2.Rmd, to explore how temperature could have affected seagrass/epifauna response variables. Going to use hourly temperature data from both the harbor (Jess Abbott's hobologgers at Doran and Westside eelgrass) and from the BOON observing sensors (shorestation and buoy)
Also, just nota bene: I compiled all sorts of BOON data into one massive dataset here: compiled.BOON.csv
The original code can be found in organizingBOON.R

# Temperature: do a regression between BOON and compiled jess/laura temp data to model harbor data; apply modeled temp to seagrass/epifauna response vars
-House-keeping -- working directory and necessary packages
-Import survey data, jess data, and BOON data
-Clean all the temperature time series data to be standardized
```{r packages}
library(tidyr)
library(dplyr)
library(ggplot2)
library(lubridate)
library(car)
library(gvlma)
library(mgcv)
library(nlme)
library(multcomp)
library(bbmle)
```
```{r Importing and cleaning hourly boon data}
setwd("~/Google Drive/Grace Ha Projects/Bodega Harbor/Abundance Surveys/Data")

# hourly shorestation data from 2011-2014
df <- read.csv("bml_seawater_temperature_2011_hourly.csv", header=FALSE)
sh.sal.h11 <- df[, c(2,4)] %>% 
  rename(datetime=V2, shorestation.temp=V4) %>%
  unique()
df <- read.csv("bml_seawater_temperature_2012_hourly.csv", header=FALSE)
sh.sal.h12 <- df[, c(2,4)] %>% 
  rename(datetime=V2, shorestation.temp=V4) %>%
  unique()
df <- read.csv("bml_seawater_temperature_2013_hourly.csv", header=FALSE)
sh.sal.h13 <- df[, c(2,4)] %>% 
  rename(datetime=V2,shorestation.temp=V4) %>%
  unique()
df <- read.csv("bml_seawater_temperature_2014_hourly.csv", header=FALSE)
sh.sal.h14 <- df[, c(2,4)] %>% 
  rename(datetime=V2, shorestation.temp=V4) %>%
  unique()

sh1 <- full_join(sh.sal.h11, sh.sal.h12)
sh2 <- full_join(sh.sal.h13, sh.sal.h14)
sh3 <- full_join(sh1, sh2)
sh3$datetime <- ymd_hms(sh3$datetime)
sh3$datetime <- with_tz(sh3$datetime, "America/Los_Angeles")



# hourly buoy data from 2012-2014
df <- read.csv("buoy_wqm_temperature_2012_hourly.csv", header=FALSE)
buoy.temp.h12 <- df[, c(2,4)] %>% 
  rename(datetime=V2, buoy.temp=V4) %>%
  unique()
df <- read.csv("buoy_wqm_temperature_2013_hourly.csv", header=FALSE)
buoy.temp.h13 <- df[, c(2,4)] %>% 
  rename(datetime=V2,  buoy.temp=V4) %>%
  unique()
df <- read.csv("buoy_wqm_temperature_2014_hourly.csv", header=FALSE)
buoy.temp.h14 <- df[, c(2,4)] %>% 
  rename(datetime=V2,  buoy.temp=V4) %>%
  unique()

buoy1 <- full_join(buoy.temp.h12, buoy.temp.h13)
buoy2 <- full_join(buoy1, buoy.temp.h14)
buoy2$datetime <- ymd_hms(buoy2$datetime)
buoy2$datetime <- with_tz(buoy2$datetime, "America/Los_Angeles")

hr <- full_join(sh3, buoy2) 

# Jess's intertidal hobologger data from doran and westside
jess.doran <- read.csv("jess_doran_temp.csv", header=TRUE) %>%
    mutate(doran.temp = (temp.f - 32)*(5/9)) %>%
    rename(datetime = datetime.gmtmin7) %>%
    select_at(vars(datetime, doran.temp))

jess.westside <- read.csv("jess_westside_temp.csv", header=TRUE) %>%
    mutate(westside.temp = (temp.f - 32)*(5/9)) %>%
    rename(datetime = datetime.gmtmin7) %>%
    select_at(vars(datetime, westside.temp))

jess <- full_join(jess.doran, jess.westside)

jess.westside$datetime <- as.POSIXct(jess.westside$datetime, format = "%m/%d/%y %H:%M" , tz = "America/Los_Angeles")
jess.doran$datetime <- as.POSIXct(jess.doran$datetime, format = "%m/%d/%y %H:%M" , tz = "America/Los_Angeles")
jess$datetime <- as.POSIXct(jess$datetime, format = "%m/%d/%y %H:%M" , tz = "America/Los_Angeles")

# Combining Jess (harbor) and BOON data into one dataset
temp.data <- full_join(jess, hr) 
```
```{r plotting all of the temps}
temp.melted <- temp.data %>%
    rename(doran = doran.temp, westside = westside.temp, 
           shorestation = shorestation.temp, buoy = buoy.temp) %>%
    gather(site, temp, doran, westside, shorestation)
all.temps <- ggplot(temp.melted, aes(x=datetime, y=temp)) + geom_point(aes(colour=site))
all.temps
ggsave("harbor.shore.temps.png", height = 6, width = 8, units = "in")
```
```{r diff between sites in temp?}
# just getting doran and westside
data <- temp.data %>%
    filter(is.na(doran.temp) == FALSE) %>%
    filter(is.na(westside.temp) == FALSE)

tvar <- data %>%
    gather(site, temp, doran.temp, westside.temp) 
tvar1 <- tvar %>%
    group_by(site) %>%
    summarise(tvar = var(temp)) 
with(tvar1, max(tvar)/min(tvar)) 
#greater than 3? Move on to Mann-Whitney test; if not, stay with Levene/t-test

# testing for normality
m <- lm(temp ~ site, tvar)
shapiro.test(head(sample(m$residuals), 5000)) # using a randomized pool of 5000 residuals; W is greater than 0.95
leveneTest(m) # significant levene's test --> var.equal=FALSE
plot(m) # decent residual plot, pretty good QQ plot

t.test(temp ~ site, tvar, var.equal = FALSE) 
boxplot(temp ~ site, tvar)
```
Cannot use the same regression for both sites; they are significantly different in temp


Apply regression between BOON data and jess/laura data
```{r comparing shorestation and westside}
df <- temp.data %>%
    filter(shorestation.temp > 7 & shorestation.temp < 18 & 
               is.na(shorestation.temp) == FALSE) %>%
    filter(westside.temp > 7 & westside.temp < 18 & 
               is.na(westside.temp) == FALSE) %>%
    filter(doran.temp > 7 & doran.temp < 18 & is.na(doran.temp) == FALSE) %>%
    mutate(year = year(datetime))

with(df, plot(westside.temp ~ shorestation.temp)) 
with(df[df$year == 2012,], plot(westside.temp ~ datetime))
points(shorestation.temp ~ datetime, df[df$year == 2012,], col = "red")

with(df[df$year >2012,], plot(westside.temp ~ datetime))
points(shorestation.temp ~ datetime, df[df$year >2012 ,], col = "red")
with(df[df$year > 2012,], plot(westside.temp ~ shorestation.temp)) 

temp.gam <- gam(westside.temp ~ s(shorestation.temp), data = df)
plot(temp.gam) 
```
```{r comparing shorestation and doran}
with(df, plot(doran.temp ~ shorestation.temp)) 
with(df[df$year == 2012,], plot(doran.temp ~ datetime))
points(shorestation.temp ~ datetime, df[df$year == 2012,], col = "red")

with(df[df$year >2012,], plot(doran.temp ~ datetime))
points(shorestation.temp ~ datetime, df[df$year >2012 ,], col = "red")
with(df[df$year > 2012,], plot(doran.temp ~ shorestation.temp)) 

temp.gam <- gam(doran.temp ~ s(shorestation.temp), data = df)
plot(temp.gam) 
```

```{r modeling harbor sites from shorestation}
model <- lm(westside.temp ~ shorestation.temp, df) # R^2 = 0.58
modsum <- summary(model)

intercept.wes <- coef(modsum)[1]
slope.wes <- coef(modsum)[2]
# --> westside.pred = slope.wes*shorestation + intercept.wes

model <- lm(doran.temp ~ shorestation.temp, df) # R^2 = 0.36
modsum <- summary(model)

intercept.dor <- coef(modsum)[1]
slope.dor <- coef(modsum)[2]
# --> doran.pred = slope.dor*shorestation + intercept.dor

temp.df1 <- temp.data %>%
    select_at(vars(datetime, shorestation.temp)) %>%
    filter(is.na(shorestation.temp) == FALSE) %>%
    mutate(dorantemp.pred = slope.dor*shorestation.temp + intercept.dor,
        westsidetemp.pred = slope.wes*shorestation.temp + intercept.wes, 
        month = month(datetime), year = year(datetime)) %>%
    filter(year > 2012) %>%
    group_by(year, month) %>%
    summarize_at(vars(dorantemp.pred, westsidetemp.pred), funs(mean))

temp.df2 <- as.data.frame(temp.df1[9:24,])

write.csv(temp.df2, "predicted.temp.csv")
```

